\LARGE{{\bf{Gravitational Wave Signals from Multiple Hidden Sectors} \\ }}

### 2019-10-04 gwnnDraft02.tex
even smaller nitpicks
abstract
changes in sec 6

The exotic sectors have essentially massless degrees of freedom pre-phase transition and pions with negligible masses post-phase transition. Other composite particles, such as baryons, do gain a mass of the order of $\Lambda_{ex}$; this is, however, still much smaller than the order of $\rho_R$ leading to small $\alpha_{\infty}$ according to Eq.~(\ref{eqn:critPTstrength}).\footnote {It should also be noted that although free quarks cease to exist post phase transition in these exotic sectors, their masses are so light that they do not contribute relevant amounts to $\alpha_{\infty}$.} Heavy standard sectors that undergo SFOPT for QCD feature no baryons due to all quarks being above their respective QCD scales. They do, however, feature glueballs that obtain a mass of the order of the SFOPT and, just as in the case for exotic sectors above, feature small $\alpha_{\infty}$. Ultimately, calculating the critical phase transition strength for each sector using Eq.~(\ref{eqn:critPTstrength}) indicates that every sector has a phase transition in the runaway regime. Explicitly, all exotic sectors (and heavy standard sectors) have strong first order phase transitions that lead to runaway bubble walls ($\alpha >\alpha_{\infty}$). Each component of the spectral energy density in Eq.~(\ref{eqn:SpectralEng}) is proportional to a power of their respective efficiency factors $\kappa$. In this case, the efficiency factors for the sound wave and MHD contributions are small and the GWs produced are dominantly from bubble collisions, $h^2\Omega_{\textrm{GW}} \approx h^2\Omega_{\phi}$.

@@

Now that the framework has been laid out for the creation of GW from a single SFOPT, we generalize to multiple sectors going under independent, coherent SFOPT. In the models presented in this paper, we consider a subset of $N$ hidden sectors that undergo a phase transition at a SM temperature of $T_{\gamma}^{i}$. As the GWs propagate in free space, the energy density and frequency spectrum, at the time of production $\Omega_{\textrm{GW}}^{*}(f)$, will redshift to today's value $\Omega_{\textrm{GW}}^{0}(f) = \mathcal{A}\, \Omega_{\textrm{GW}}^{*}((a_{0}/a) f) $. The redshifting factor, $\mathcal{A}$, accounts for the redshifting of both $\rho_{\textrm{GW}} $ and $\rho_{c}$~\cite{Breitbach:2018ddu,PhysRevD.49.2837},
\begin{equation}\label{eqn::redshifting}
\mathcal{A} \equiv \bigg( \frac{a}{a_{0}} \bigg)^4 \bigg( \frac{H}{H_{0}}\bigg)^2
\end{equation}
where $a$ ($a_{0}$)  and $H$ ($H_{0}$) are the scale factor and Hubble constant at the time of nucleation (observation), respectively. Assuming that the sectors are completely decoupled before and after their respective SFOPT, the total GW signal that would be measured today is given by the coherent sum,
\begin{equation}
\Omega_{\textrm{GW}} =  \sum_{i}^{N} \mathcal{A}^{i}\, \Omega_{\textrm{GW}}^{i,*}((a_{0}/a)_{i} f) .
\end{equation}
We assume that the parameters of the SFOPT do not differ between sectors: the relativistic degrees of freedom, phase transition rate, and the dark QCD scale, are all similar. This makes the redshifting factor $\mathcal{A}^i$ independent of sector number.
\begin{figure}[h!]
\centering
\includegraphics[scale=.3]{Nnatral.png} 
\caption{ Gravitational wave spectral energy density (solid curves) for standard $N$naturalness using the scalar reheaton model of section~\ref{sec:reheat}. The curve corresponding to the sum of the sectors is approximately equal to the $i=1$ curve. All contributions are assumed to be purely from bubble collisions $\Omega_{\phi}$, with $\beta/H  = 10$.  The shaded dashed curves are the power law noise curves~\cite{PhysRevD.88.124032} calculated from expected sensitivity as described in Section~\ref{sec:detection}. The ones on the right are space-based interferometers: Lisa~\citep{Audley:2017drz} (blue), DECIGO~\citep{10.1093/ptep/pty078} (light blue), BBO~\citep{PhysRevD.72.083005} (red). The ones on the left are for the pulsar timing array SKA~\cite{Janssen:2014dka} for exposure time of 5-year (purple), 10-year (orange), and 20-year (green). }
\label{fig::Nnatural}
\end{figure}
Applying this to the standard reheating scenario of $N$natrualness, introduced in~\ref{sec:reheat}, we get GW signals as seen in Fig.~\ref{fig::Nnatural}. Plotted are the individual contributions to the signal from each phase transitioned sector, as well as the coherent sum of all sectors. Future GW interferometers and pulsar timing array sensitivity curves are shown in comparison to the signal. The sensitivity curves are interpreted as the region of possible detection if intersected with the GW signal, more detail in the construction of these curves is given in Section~\ref{sec:detection}. Notice that the total signal is dominated by the first sectors contribution. This is caused by the quartic temperature ratio suppression in Eq.~(\ref{eq:alpha}) and the large temperature gaps between adjacent sectors. Such a suppression leads to standard $N$natrualness evading future detector thresholds by a few orders of magnitude in units of energy density.

This is not the case if we consider more generalized reheating scenarios. Once the restriction that sectors with small Higgs masses are preferentially reheated has been lifted, we can explore a much more vast landscape of hidden sectors than are allowed in the vanilla case. Here, we construct several different scenarios that are both detectable and demonstrate a variety of gravitational wave profiles. Specifically, we explore benchmarks that lead to a deviation in the peak behaviour of the total GW signal (the superposition of stochastic GW from individual SFOPT) from a standard power law signal.

@@

In the following, we focus on heavy standard sectors --- pure Yang-Mills sectors with much heavier particles (specifically quarks) and, as shown in Sec.~\ref{sec:dQCD}, the SFOPT these entail. The reason for this arises from Eq.~(\ref{eqn:peakFrequency}): every exotic sector features a phase transition that occurs at $\Lambda_{ex} \sim 90$ MeV. If we maximize the allowed temperature ratio, this gives us a (SM) photon temperature, $T_{\gamma}$, that places our signal directly in the frequency void between the detection region of pulsar timing arrays and space-based interferometers (see Sec.~\ref{sec:detection}). The location of the peak can be changed by dropping the temperature ratio, but the adjustment required to end up with a signal with an appropriate peak frequency makes the overall signal too weak to detect. As shown in Sec.~\ref{sec:dQCD}, standard sectors can have much higher temperature phase transitions. As such, maintaining the maximum allowed temperature ratio between the hidden sector(s) and the SM gives a much larger photon temperature and a proportionally larger peak frequency; ultimately allowing for detection by space-based interferometers. There are four scenarios that we examine, with key parameters presented in Tab.~\ref{tab:cases}.

\begin{table*}[t]
\begin{center}
\begin{tabular}{ |p{2cm}||p{3cm}|p{3cm}|p{3cm}|p{2cm}|  }
 \hline
 \multicolumn{5}{|c|}{Parameters for multi-hidden sector benchmarks} \\
 \hline
\multicolumn{5}{|c|}{Maximized signal} \\
 \hline
 Sector & Higgs vev (GeV) & $\Lambda^{AS}_{QCD}$ (GeV) & $T_\gamma$ (GeV) & Index\\
 \hline
 1 & $24.6 \times 10^{9}$ & 38.6 & 87.7 & $10^{16}$ \\
 \hline
\multicolumn{5}{|c|}{Large split} \\
 \hline
 Sector & Higgs vev (GeV) & $\Lambda^{AS}_{QCD}$ (GeV) & $T_\gamma$ (GeV) & Index\\
 \hline
 1 & $246 \times 10^6$  & 10.8 & 30.3 & $10^{12}$ \\
 2 & $7.8 \times 10^9$  & 28.1 & 78.6 & $10^{15}$ \\
 \hline
 \multicolumn{5}{|c|}{Medium split} \\
 \hline
 Sector & Higgs vev (GeV) & $\Lambda^{AS}_{QCD}$ (GeV) & $T_\gamma$ (GeV) & Index\\
 \hline
 1 & $246 \times 10^{6}$ & 10.8 & 30.3 & $10^{12}$ \\
 2 & $778 \times 10^6$  & 14.9 & 41.6 & $10^{13}$ \\
 \hline
 \multicolumn{5}{|c|}{Five sector} \\
 \hline
 Sector & Higgs vev (GeV) & $\Lambda^{AS}_{QCD}$ (GeV) & $T_\gamma$ (GeV) & Index\\
 \hline
 1 & $246 \times 10^6$ & 10.8 & 38.8 & $10^{12}$ \\
 2 & $426 \times 10^6$ & 12.6 & 45.2 & $3\times10^{12}$ \\
 3 & $778 \times 10^6$ & 14.9 & 53.3 & $10^{13}$ \\
 4 & $1.3 \times 10^9$ & 17.3 & 62.1 & $3\times10^{13}$ \\
 5 & $2.5 \times 10^9$ & 20.5 & 73.3 & $10^{14}$ \\
 \hline
\end{tabular}
\caption{Outline of parameters used for the various multi-hidden sector scenarios. The higgs vev is the vev for the given additional sector, $\Lambda^{AS}_{QCD}$ is the QCD phase transition in the additional sector, and $T_\gamma$ is the temperature of the SM photon bath when the SFOPT occurs in the additional sector. The index indicates the equivalent sector from the $N$naturalness model (Eq.~(\ref{eqn:vevs})). It should be noted that although the various sectors transition at different temperatures, they are all assumed to be reheated to the same initial temperature as discussed in Sec.~\ref{sec:genReheat}.}\label{tab:cases}
\end{center}
\end{table*}

\begin{itemize}
\item \textbf{Maximized signal:} A single additional heavy hidden sector reheated to a temperature that saturates current experimental bounds. The SM photon bath temperature at the time of the hidden sector PT is $87$ GeV. In the $N$naturalness framework this is equivalent to reheating a standard sector with $i\sim 10^{16}$ up to the maximum allowed temperature ratio.

@@

\end{itemize}

>>>>

The exotic sectors have essentially massless degrees of freedom pre-phase transition and pions with negligible masses post-phase transition. Other composite particles, such as baryons, do gain a mass of the order of $\Lambda_{ex}$; this is, however, still much smaller than the order of $\rho_R$ leading to small $\alpha_{\infty}$ according to Eq.~(\ref{eqn:critPTstrength}).\footnote {It should also be noted that although free quarks cease to exist post phase transition in these exotic sectors, their masses are so light that they do not contribute relevant amounts to $\alpha_{\infty}$.} Heavy standard sectors that undergo SFOPT for QCD feature no baryons due to all quarks being above their respective QCD scales. They do, however, feature glueballs that obtain a mass of the order of the SFOPT and, just as in the case for exotic sectors above, feature small $\alpha_{\infty}$. Ultimately, calculating the critical phase transition strength for each sector using Eq.~(\ref{eqn:critPTstrength}) indicates that every sector has a phase transition in the runaway regime. 
%Explicitly, all exotic sectors (and heavy standard sectors) have strong first order phase transitions that lead to runaway bubble walls ($\alpha >\alpha_{\infty}$). 
Each component of the spectral energy density in Eq.~(\ref{eqn:SpectralEng}) is proportional to a power of their respective efficiency factors $\kappa$. In this case, the efficiency factors for the sound wave and MHD contributions are small and the GWs produced are dominantly from bubble collisions, $h^2\Omega_{\textrm{GW}} \approx h^2\Omega_{\phi}$.

@@

Now that the framework has been laid out for the creation of GW from a single SFOPT, we generalize to multiple sectors going under independent, coherent SFOPT. In the models presented in this paper, we consider a subset of $N$ hidden sectors that undergo a phase transition at a SM temperature of $T_{\gamma}^{i}$. As the GWs propagate in free space, the energy density and frequency spectrum, at the time of production $\Omega_{\textrm{GW}}^{*}(f)$, will redshift to today's value $\Omega_{\textrm{GW}}^{0}(f) = \mathcal{A}\, \Omega_{\textrm{GW}}^{*}((a_{0}/a) f) $. The redshifting factor, $\mathcal{A}$, accounts for the redshifting of both $\rho_{\textrm{GW}} $ and $\rho_{c}$~\cite{Breitbach:2018ddu,PhysRevD.49.2837},
\begin{equation}\label{eqn::redshifting}
\mathcal{A} \equiv \bigg( \frac{a}{a_{0}} \bigg)^4 \bigg( \frac{H}{H_{0}}\bigg)^2
\end{equation}
where $a$ ($a_{0}$)  and $H$ ($H_{0}$) are the scale factor and Hubble constant at the time of nucleation (observation), respectively. Assuming that the sectors are completely decoupled before and after their respective SFOPT, the total GW signal that would be measured today is given by the coherent sum,
\begin{equation}
\Omega_{\textrm{GW}} =  \sum_{i}^{N} \mathcal{A}^{i}\, \Omega_{\textrm{GW}}^{i,*}((a_{0}/a)_{i} f) .
\end{equation}
We assume that the parameters of the SFOPT do not differ between sectors: the relativistic degrees of freedom, phase transition rate, and the dark QCD scale, are all similar. This makes the redshifting factor $\mathcal{A}^i$ independent of sector number.
\begin{figure}[h!]
\centering
\includegraphics[scale=.3]{Nnatral.png} 
\caption{ Gravitational wave spectral energy density (solid curves) for standard $N$naturalness using the scalar reheaton model of section~\ref{sec:reheat}. The curve corresponding to the sum of the sectors is approximately equal to the $i=1$ curve. All contributions are assumed to be purely from bubble collisions $\Omega_{\phi}$, with $\beta/H  = 10$.  The shaded dashed curves are the power law noise curves~\cite{PhysRevD.88.124032} calculated from expected sensitivity as described in Section~\ref{sec:detection}. The ones on the right are space-based interferometers: Lisa~\citep{Audley:2017drz} (blue), DECIGO~\citep{10.1093/ptep/pty078} (light blue), BBO~\citep{PhysRevD.72.083005} (red). The ones on the left are for the pulsar timing array SKA~\cite{Janssen:2014dka} for exposure time of 5-years (purple), 10-years (orange), and 20-years (green). }
\label{fig::Nnatural}
\end{figure}
Applying this to the standard reheating scenario of $N$natrualness, introduced in~\ref{sec:reheat}, we get GW signals as seen in Fig.~\ref{fig::Nnatural}. Plotted are the individual contributions to the signal from each phase transitioned sector, as well as the coherent sum of all sectors. Future GW interferometers and pulsar timing array sensitivity curves are shown in comparison to the signal. The sensitivity curves are interpreted as the region of possible detection if intersected with the GW signal, and the construction of these curves is detailed in Section~\ref{sec:detection}. Notice that the total signal is dominated by the first sector's contribution. This is caused by the quartic temperature ratio suppression in Eq.~(\ref{eq:alpha}) and the large temperature gaps between adjacent sectors. Such a suppression leads to standard $N$natrualness evading future detector thresholds by a few orders of magnitude in units of energy density.

This is not the case if we consider more generalized reheating scenarios. Once the restriction that sectors with small Higgs masses are preferentially reheated has been lifted, we can explore a much more vast landscape of hidden sectors than are allowed in the reheaton case. Here, we construct several different scenarios that are both detectable and demonstrate a variety of gravitational wave profiles. Specifically, we explore benchmarks that lead to a deviation in the peak behaviour of the total GW signal (the superposition of stochastic GW from individual SFOPT) from a standard power law signal.

@@

In the following, we focus on heavy standard sectors --- pure Yang-Mills sectors with much heavier particles (specifically quarks) and, as shown in Sec.~\ref{sec:dQCD}, the SFOPT these entail. The reason for this arises from Eq.~(\ref{eqn:peakFrequency}): every exotic sector features a phase transition that occurs at $\Lambda_{ex} \sim 90$ MeV. If we maximize the allowed temperature ratio, this gives us a (SM) photon temperature, $T_{\gamma}$, that places our signal directly in the frequency void between the detection region of pulsar timing arrays and space-based interferometers (see Sec.~\ref{sec:detection}). The location of the peak can be changed by dropping the temperature ratio, but the adjustment required to end up with a signal with an appropriate peak frequency makes the overall signal too weak to detect. As shown in Sec.~\ref{sec:dQCD}, standard sectors can have much higher temperature phase transitions. As such, maintaining the maximum allowed temperature ratio between the hidden sector(s) and the SM gives a much larger photon temperature and a proportionally larger peak frequency; ultimately allowing for detection by space-based interferometers.

\begin{table*}[t]
\begin{center}
\begin{tabular}{ |p{2cm}||p{3cm}|p{3cm}|p{3cm}|p{2cm}|  }
 \hline
 \multicolumn{5}{|c|}{Parameters for multi-hidden sector benchmarks} \\
 \hline
\multicolumn{5}{|c|}{Maximized signal} \\
 \hline
 Sector & Higgs vev (GeV) & $\Lambda^{AS}_{QCD}$ (GeV) & $T_\gamma$ (GeV) & Index\\
 \hline
 1 & $24.6 \times 10^{9}$ & 38.6 & 87.7 & $10^{16}$ \\
 \hline
\multicolumn{5}{|c|}{Large split} \\
 \hline
 Sector & Higgs vev (GeV) & $\Lambda^{AS}_{QCD}$ (GeV) & $T_\gamma$ (GeV) & Index\\
 \hline
 1 & $246 \times 10^6$  & 10.8 & 30.3 & $10^{12}$ \\
 2 & $7.8 \times 10^9$  & 28.1 & 78.6 & $10^{15}$ \\
 \hline
 \multicolumn{5}{|c|}{Medium split} \\
 \hline
 Sector & Higgs vev (GeV) & $\Lambda^{AS}_{QCD}$ (GeV) & $T_\gamma$ (GeV) & Index\\
 \hline
 1 & $246 \times 10^{6}$ & 10.8 & 30.3 & $10^{12}$ \\
 2 & $778 \times 10^6$  & 14.9 & 41.6 & $10^{13}$ \\
 \hline
 \multicolumn{5}{|c|}{Five sector} \\
 \hline
 Sector & Higgs vev (GeV) & $\Lambda^{AS}_{QCD}$ (GeV) & $T_\gamma$ (GeV) & Index\\
 \hline
 1 & $246 \times 10^6$ & 10.8 & 38.8 & $10^{12}$ \\
 2 & $426 \times 10^6$ & 12.6 & 45.2 & $3\times10^{12}$ \\
 3 & $778 \times 10^6$ & 14.9 & 53.3 & $10^{13}$ \\
 4 & $1.3 \times 10^9$ & 17.3 & 62.1 & $3\times10^{13}$ \\
 5 & $2.5 \times 10^9$ & 20.5 & 73.3 & $10^{14}$ \\
 \hline
\end{tabular}
\caption{Outline of parameters used for the various multi-hidden sector scenarios. The higgs vev is the vev for the given additional sector, $\Lambda^{AS}_{QCD}$ is the QCD phase transition in the additional sector, and $T_\gamma$ is the temperature of the SM photon bath when the SFOPT occurs in the additional sector. The index indicates the equivalent sector from the $N$naturalness model (Eq.~(\ref{eqn:vevs})). It should be noted that although the various sectors undergo phase transitions at different temperatures, they are all assumed to be reheated to the same initial temperature.}\label{tab:cases}
\end{center}
\end{table*}

There are four scenarios that we examine, with key parameters presented in Tab.~\ref{tab:cases}. 
\begin{itemize}
\item \textbf{Maximized signal:} A single additional heavy hidden sector reheated to a temperature that saturates current experimental bounds. The SM photon bath temperature at the time of the hidden sector PT is $87$ GeV. In the $N$naturalness framework this is equivalent to reheating a standard sector with $i\sim 10^{16}$ up to the maximum allowed temperature ratio.

@@

\end{itemize}
%
In all cases where multiple sectors are reheated, we assume for simplicity that all the hidden sectors are reheated to the same temperature.
